---
title: "R Notebook"
output: 
  html_notebook:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

In this document, I will create a graph that shows the distribution of MLGs
across populations

```{r}
library('tidyverse')
library('assertr')
library('poppr')
library('igraph')
library("ggraph")
library('magrittr')
```

## Loading data and assertions


```{r read_data}
dat <- read.genalex("../Analysis4 ForManu/A2_Copy4 EUR_AUS_forManu.csv", ploidy = 1)
splitStrata(dat) <- ~Isolate/Severity/MCG/Region/Source/Year/Host
dat

repeat_lengths <-
  c(
  `5-2` = 2.000000,
  `5-3` = 0.500000,
  `6-2` = 6.000000,
  `7-2` = 2.000000,
  `8-3` = 2.000000,
  `9-2` = 2.000000,
  `12-2` = 2.000000,
  `17-3` = 3.000000,
  `20-3` = 2.000000,
  `36-4` = 0.500000,
  `50-4` = 0.500000,
  `55-4` = 4.000000,
  `92-4` = 2.000000,
  `106-4` = 4.000000,
  `110-4` = 4.000000,
  `114-4` = 4.000000
  )
repeat_lengths <- ifelse(repeat_lengths < 1, 4, repeat_lengths)
```


```{r excel_data}
ex <- readxl::read_excel("../Analysis4 ForManu/A1_Copy of binned-genotypes_SE.xlsx", sheet = "GenAlex", skip = 1) %>%
  select(-1) %>%                # removing first column, which is empty
  gather(locus, allele, -1) %>% # gather all loci into tidy columns
  mutate(locus = trimws(locus)) %>% # remove (F) designator
  mutate(allele = as.integer(allele)) %>% # force alleles to integers
  spread(locus, allele) %>%
  slice(-n())
ex <- ex[!names(ex) %in% locNames(dat)]

# Function to select an adjacent allele. It will select the
# next allele if the next allele is not missing and it's distance
# is one away and the previous allele for the same conditions.
# If none of the conditions are met, it will retain the allele.
cromulent_allele <- Vectorize(function(lower, allele, higher){
  if (!is.na(higher) && abs(allele - higher) == 1){
    out <- higher
  } else if (!is.na(lower) && abs(allele - lower) == 1){
    out <- lower
  } else {
    out <- allele
  }
  out
})
ex
exsummary <- ex %>% 
  gather(locus, allele, -1) %>% # tidy the data
  group_by(locus, allele) %>%   
  summarize(n = n()) %>%        # summarize by count 
  ungroup() %>%
  group_by(locus) %>%           # group the loci, add the lower and upper alleles,
  mutate(lower = lag(allele), higher = lead(allele)) %>% # and then create new_alleles
  mutate(new_allele = ifelse(n < 3, cromulent_allele(lower, allele, higher), allele)) %>%
  select(locus, new_allele, allele)
exsummary
corrected_loci <- ex %>% gather(locus, allele, -1) %>%
  left_join(exsummary, by = c("locus", "allele")) %>%
  mutate(allele = new_allele) %>%
  select(-new_allele) %>%
  spread(locus, allele)
datdf <- genind2df(dat, usepop = FALSE) %>% bind_cols(corrected_loci[-1])
dat  <- df2genind(datdf, ind.names = indNames(dat), strata = strata(dat), ploidy = 1) %>% as.genclone()
```

The incoming strata includes both Severity and Isolate. Since these are not
necessary for delimiting the strata, we will place them in the "other" slot
after converting Severity to numeric. Placing this information in the "other"
slot ensures that these data will travel with the object.

```{r fix_strata}
dat_strata <- strata(dat) %>%
  mutate_all(as.character) %>%
  mutate(Severity = as.numeric(Severity))
strata(dat)     <- select(dat_strata, -Severity, -Isolate)
indNames(dat)   <- dat_strata$Isolate
other(dat)$meta <- select(dat_strata, Severity, Isolate)
```

```{r}
setPop(dat) <- ~Region
dat
keeploci <- !locNames(dat) %in% colnames(corrected_loci)
genotype_curve(dat, sample = 1000)
locus_table(dat)
# rl <- fix_replen(dat, repeat_lengths)
# mlg.filter(dat, distance = bruvo.dist, replen = rl) <- .Machine$double.eps
dat
genotype_curve(dat[loc = keeploci, mlg.reset = TRUE])
# dat <- dat[loc = keeploci, mlg.reset = TRUE]
dat
```

## Crossing populations


We can use `mlg.crosspop()` to tabulte which MLGs cross populations.

I realized that it's possible to use an MLG table with matrix multiplication to
get an adjency matrix.

```{r}
make_graph_list <- function(dat){ # dat is a genclone/snpclone object
  datmlg  <- mlg.table(dat, plot = FALSE) > 0 # presence/absence of MLG
  crosses <- mlg.crosspop(dat, quiet = TRUE, df = TRUE) %>% tbl_df()
  adjmat  <- datmlg %*% t(datmlg) 
  cols    <- sort(colnames(adjmat))
  adjmat  <- adjmat[cols, cols]
  # Creating Graph
  g           <- graph_from_adjacency_matrix(adjmat, mode = "undirected", diag = FALSE)
  V(g)$size   <- diag(adjmat)
  g           <- delete_vertices(g, degree(g) == 0)
  shared_mlg  <- (crosses %>% group_by(Population) %>% summarize(n = n()))$n
  V(g)$weight <- 1 - shared_mlg/V(g)$size
  el          <- as_adj_edge_list(g)
  el          <- el[lengths(el) > 0]
  popgraphs <- setNames(vector(mode = "list", length = length(el) + 1), c(names(el), "total"))
  for (v in names(el)){
    idx  <- el[[v]]
    mlgs <- crosses %>%           # How to get all MLGs from a single population:
      filter(Population == v) %>%         # Grab only the population e and then
      select(MLG) %>%                     # remove everything but the MLGs to do an
      inner_join(crosses, by = "MLG") %>% # inner join of the original list and then
      filter(Population != v) %>%         # remove the query population to give
      arrange(Population)                 # the neigboring populations in order.
    MLGS <- as.character(mlgs$MLG)
    E(g)[idx]$label  <- substr(MLGS, 5, nchar(MLGS))
    E(g)[idx]$weight <- as.integer(table(MLGS)[MLGS]) # weight == n populations visited
    popgraphs[[v]]   <- subgraph.edges(g, eids = idx)
  }
  popgraphs[["total"]] <- g
  popgraphs
}

plot_mlg_graph <- function(g, glayout = NULL){
  if (is.null(glayout)){
    glayout <- layout_nicely(g)
  } else {
    glayout <- glayout[V(g)$name, ]
  }
  shared_mlg  <- (1 - V(g)$weight) * V(g)$size
  g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
  g2          <- g
  V(g2)$label <- ifelse(!is.na(V(g2)$name), sprintf("%s [(%d/%d)]", V(g2)$name, shared_mlg, V(g2)$size), NA)
  glay        <- create_layout(g2, "manual", node.positions = as.data.frame(rbind(glayout, glayout)))
  x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
  ggraph(glay) +
    geom_edge_fan(aes(alpha = weight + 1), width = 0.75) +
    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size)) +
    geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) +
    viridis::scale_fill_viridis(option = "C") +
    scale_edge_alpha_continuous(range = c(0.25, 1), breaks = c(2:5)) +
    coord_fixed() +
    theme_void() +
    theme(text = element_text(size = 14)) +
    labs(list(
      title = "Shared haplotypes across regions",
      fill = "Number of\nGenotypes",
      edge_alpha = "Populations\nper haploytpe",
      caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
    ))
}

plot_mlg_subgraph <- function(graphlist){
  for (i in names(graphlist)){
    pg   <- graphlist[[i]]
    labs <- ifelse(E(pg)$weight > 1, E(pg)$label, NA)
    labs <- ifelse(duplicated(labs), NA, labs)
    plot(pg, 
         main = i, 
         layout = layout_as_star(pg, center = i), 
         edge.width = E(pg)$weight,
         edge.label = labs)
  }
}



good_layout <- read.table(
text = 
"   x                            y
AU  0.8090169943749470071737  5.877852522924730260812e-01
CA -1.0000000000000000000000  1.224646799147350002426e-16
CO -0.3090169943749480063744 -9.510565162951539752711e-01
FR -0.3090169943749470071737  9.510565162951539752711e-01
MI  0.0000000000000000000000  0.000000000000000000000e+00
MN  1.0000000000000000000000  0.000000000000000000000e+00
ND  0.3090169943749470071737 -9.510565162951539752711e-01
NE -0.8090169943749470071737 -5.877852522924730260812e-01
NY  0.8090169943749470071737 -5.877852522924730260812e-01
OR  0.3090169943749470071737  9.510565162951539752711e-01
WA -0.8090169943749470071737  5.877852522924730260812e-01"
) %>% as.matrix()

```


## Graphs

With the fuctions above, we can create and plot the graphs

```{r, fig.width = 7, fig.height = 7}
# Creating the graphs
graph16loc <- make_graph_list(dat)
graph11loc <- make_graph_list(dat[loc = keeploci, mlg.reset = TRUE])

# Plotting the subgraphs
par(mfrow = c(3, 4))
plot_mlg_subgraph(graph16loc[-length(graph16loc)])

par(mfrow = c(3, 4))
plot_mlg_subgraph(graph11loc[-length(graph11loc)])

# Plotting the full graphs
par(mfrow = c(1, 1))
plot_mlg_graph(graph16loc$total, good_layout)
plot_mlg_graph(graph11loc$total, good_layout)
```

```{r, eval = FALSE}
null_vertices <- length(V(g)) %>% seq(from = ./2 + 1, to = ., by = 1)
sg <- g %>% 
  delete_vertices(null_vertices) %>%
  cluster_spinglass()
modularity(sg)
communities(sg)
```

