%% Submissions for peer-review must enable line-numbering
%% using the lineno option in the \documentclass command.
%%
%% Preprints and camera-ready submissions do not need
%% line numbers, and should have this option removed.
%%
%% Please note that the line numbering option requires
%% version 1.1 or newer of the wlpeerj.cls file, and
%% the corresponding author info requires v1.2

\documentclass[fleqn,10pt,lineno]{wlpeerj} % for journal submissions

% ZNK -- Adding headers for pandoc

\setlength{\emergencystretch}{3em}
\providecommand{\tightlist}{
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{lipsum}
\usepackage[unicode=true]{hyperref}
\usepackage{longtable}



\usepackage{indentfirst}

\title{Population structure and phenotypic variation of \emph{Sclerotinia
sclerotiorum} from dry bean (\emph{Phaseolus vulgaris}) in the United
States}

\author[1]{Zhian N. Kamvar}

\author[1]{Bimal Sajeewa Amaradasa}

\author[1]{Rachana Jhala}

\author[1]{Serena McCoy}

\author[1]{James R. Steadman}

\author[1]{Sydney E. Everhart}

\corrauthor[1]{Sydney E. Everhart}{\href{mailto:everhart@unl.edu}{\nolinkurl{everhart@unl.edu}}}

\affil[1]{Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln,
NE 68583}


%
% \author[1]{First Author}
% \author[2]{Second Author}
% \affil[1]{Address of first author}
% \affil[2]{Address of second author}
% \corrauthor[1]{First Author}{f.author@email.com}

% 

\begin{abstract}
The ascomycete pathogen \emph{Sclerotinia sclerotiorum} is a
necrotrophic pathogen on over 400 known host plants, and is the causal
agent of white mold on dry bean. Currently, there are no known cultivars
of dry bean with complete resistance to white mold. For more than 20
years, bean breeders have been using white mold screening nurseries with
natural populations of \emph{S. sclerotiorum} to screen new cultivars
for resistance. It is thus important to know if the genetic diversity in
populations of \emph{S. sclerotiorum} within these nurseries a) reflect
the genetic diversity of the populations in the surrounding region and
b) are stable over time. Furthermore, previous studies have investigated
the correlation between mycelial compatibility groups (MCG) and
multilocus haplotypes (MLH), but none have formally tested these
patterns. We genotyped 366 isolates of \emph{S. sclerotiorum} from
producer fields and white mold screening nurseries surveyed over 10
years in 2003--2012 representing 11 states in the United States of
America, Australia, France, and Mexico at 11 microsatellite loci
resulting in 165 MLHs. Populations were loosely structured over space
and time based on analysis of molecular variance and discriminant
analysis of principal components, but not by cultivar, aggressiveness,
or field source. Of all the regions tested, only Mexico (n=18) shared no
MLHs with any other region. Using a bipartite network-based approach, we
found no evidence that the MCGs accurately represent MLHs. Our study
suggests that breeders should continue to test dry bean lines in several
white mold screening nurseries across the US to account for both the
phenotypic and genotypic variation that exists across regions.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\input{paper}
\newpage
\input{supplementary}
\end{document}
